Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities

### 2020-12-17 docs/manuscript/main.tex
merge fig4andfig5 into fig3 for rps

For the creation of a recurrence plot based on time series 
$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
reconstruction with uniform time-delay embedding 
$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
The recurrence plot is therefore a two-dimensional $N \times N$ square 
matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
whenever $X(i)$ is sufficiently close to $X(j)$: 
%%********************************[EQUATION]************************************
\begin{equation}
\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
\end{equation}
%%********************************[EQUATION]************************************
where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
RP is also characterised with a line of identity (LOI) which is a  
diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
    \caption{
	{\bf Recurrence Plots.} 
	(A) State space of the Lorenz system with controlling parameters 
	($\rho=28, \sigma=10, \beta=8/3$). 
	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
	(black circle) of a given point at $i$ is a recurrent point and is 
	represented as a black dot in the recurrence plot at location 
	$(i, j)$ or white otherwise.
	(B) Recurrence plot using the 
	three components of the Lorenz system and the RP with no embeddings 
	and threshold $\epsilon=5$.
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:mrp}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\subsection*{Structures of Recurrence Plots}
Pattern formations in the RPs can be designated either 
as topology for large-scale patterns or texture for small-scale patterns.
In the case of topology, the following pattern formations are commonly presented:
(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
uniformly distributed noise (Fig~\ref{fig:rp2}A), 
(ii) periodic and quasi-periodic systems where diagonal lines and 
checkerboard structures represent oscillating systems, e.g., sinusoidal 
signals (Fig~\ref{fig:rp2}B), 
(iii) drift where paling or darkening recurrence points away from 
the LOI is caused by drifting systems, 
e.g., logistic map (Fig~\ref{fig:rp2}C), and
(iv) disrupted where recurrence points are presented white areas or 
bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
Texture patterns in RPs can be categorised as:
(i) single or isolated recurrence points that represent rare occurring states, 
and do not persist for any time or fluctuate heavily,
(ii) dots forming diagonal lines where the length of the small-scale parallel 
lines in the diagonal are related to the ratio of determinism or predictability 
in the dynamics of the system, and
(iii) dots forming vertical and horizontal lines where the length of the 
lines represent a time length where a state does not change or change very 
slowly and these patterns formation represent discontinuities in the signal, 
and (iv) dots clustering to inscribe rectangular regions which are related 
to laminar states or singularities \cite{marwan2015}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
    \caption{
	{\bf Patterns in Recurrence Plots.} 
	Time-series with its respective recurrence plot for:
	(A) uniformly distributed noise,
	(B) super-positionet harmonic oscillation 
	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
	with a linearly increase term ($0.01*i$), and
	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:rp2}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

@@

>>>>

For the creation of a recurrence plot based on time series 
$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
reconstruction with uniform time-delay embedding 
$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
The recurrence plot is therefore a two-dimensional $N \times N$ square 
matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
whenever $X(i)$ is sufficiently close to $X(j)$: 
%%********************************[EQUATION]************************************
\begin{equation}
\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
\end{equation}
%%********************************[EQUATION]************************************
where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
(Fig~\ref{fig:rps}(A,B)) \cite{eckmann1987, marwan2007,marwan2015}.
RP is also characterised with a line of identity (LOI) which is a  
diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$.

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\subsection*{Structures of Recurrence Plots}
Pattern formations in the RPs can be designated either 
as topology for large-scale patterns or texture for small-scale patterns.
In the case of topology, the following pattern formations are commonly presented:
(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
uniformly distributed noise (Fig~\ref{fig:rps}C), 
(ii) periodic and quasi-periodic systems where diagonal lines and 
checkerboard structures represent oscillating systems, e.g., sinusoidal 
signals (Fig~\ref{fig:rps}D), 
(iii) drift where paling or darkening recurrence points away from 
the LOI is caused by drifting systems, 
e.g., logistic map (Fig~\ref{fig:rps}E), and
(iv) disrupted where recurrence points are presented white areas or 
bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
(Fig~\ref{fig:rps}F) \cite{eckmann1987, marwan2015}.
Texture patterns in RPs can be categorised as:
(i) single or isolated recurrence points that represent rare occurring states, 
and do not persist for any time or fluctuate heavily,
(ii) dots forming diagonal lines where the length of the small-scale parallel 
lines in the diagonal are related to the ratio of determinism or predictability 
in the dynamics of the system, and
(iii) dots forming vertical and horizontal lines where the length of the 
lines represent a time length where a state does not change or change very 
slowly and these patterns formation represent discontinuities in the signal, 
and (iv) dots clustering to inscribe rectangular regions which are related 
to laminar states or singularities \cite{marwan2015}.

@@

%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/rps/versions/drawing-v00.pdf}
    \caption{
	{\bf Recurrence Plots and its patterns.} 
	(A) State space of the Lorenz system with controlling parameters 
	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
	(black circle) of a given point at $i$ is a recurrent point and is 
	represented as a black dot in the recurrence plot at location 
	$(i, j)$ or white otherwise.
	(B) Recurrence plot using the 
	three components of the Lorenz system and the RP with no embeddings 
	and threshold $\epsilon=5$.
	Time-series with its respective recurrence plot for:
	(C) uniformly distributed noise,
	(D) super-positionet harmonic oscillation 
	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
	with a linearly increase term ($0.01*i$), and
	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:rps}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------
